The Spatial Relationship between Social Media, Digital Connectedness, and Social Isolation in China: Revisting Cyber-Optimism

### 2023-08-31 Social-Isolation-in-China.tex
prep for journal submission

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The Spatial Relationship between Social Media, Digital Connectedness, and Social Isolation in China: Revisting Cyber-Optimism},
  pdfauthor={Jason Gainous; Andrew W. MacDonald},
  pdfkeywords={social isolation, social media, digital
connectedness, China},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{The Spatial Relationship between Social Media, Digital
Connectedness, and Social Isolation in China: Revisting Cyber-Optimism}
\author{Jason Gainous\footnote{University of Louisville,
  jason.gainous@louisville.edu} \and Andrew W. MacDonald\footnote{Duke
  Kunshan University, andrew.macdonald@dukekunshan.edu.cn}}
\date{3/5/23}

\begin{document}
\maketitle
\begin{abstract}
Early research and populuar discourse on the promise of the internet was
optimistic. Many believed it would promote social interconnectedness
and, ultimately, be a positive force in people's lives. Since these
early predictions, research has become much less positive, usually
pointing to the negative aspects of social media. Here we argue that the
digital environment can actually create a space that deters feelings of
social isolationism especially for those living outside cities in
restrictive information environments where citizens are often
disconnected from the outside world. Using data from a nationwide survey
of Chinese citizens we conducted in 2015, we find that the more citizens
digitally connect with others, the more likely they are to move
relationships from online to offliine, and the more they use social
media the less personally isolated they feel. Futher, those living in
urban environments tend to feel more isolated, and this spatial factor
seems to moderate the aforementioned relationships. All three of those
observed relationships are stronger for those in less urban
environments.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, frame hidden, interior hidden, boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

Research is quite mixed when it comes to identifying the relationship
between social media use and feelings of isolationism. Early research
suggested that the internet would build social capital, or an
interconnectedness between people, and that social media platforms were
the ultimate way to bring people together. Since then, research
indicates that the effect may be the opposite for many. Social media may
encourage social isolation, and ultimately, social media users may feel
more disconnected from others than those who do not use social media. We
know very little about this phenomenon in China. China presents an
interesting case because the internet is bifurcated to some degree --
there is the Chinese internet and the outside internet available via
VPNs. Being isolation to the outside world may lessen feelings of
isolation, or it may increase these feelings because digital exposure to
a world is much larger than that in people's immediate environment
without the ability to have in-person connection to may make one feel
isolated. Interestingly, and somewhat ironically though, it may also
matter, too, whether someone lives in a city. Research has suggested
that urban living can actually lead to feelings of isolation.

\hypertarget{introduction-1}{%
\subsection{Introduction}\label{introduction-1}}

Research is quite mixed when it comes to identifying the relationship
between social media use and feelings of isolationism. Early research
suggested that the internet would build social capital, or an
interconnectedness between people, and that social media platforms were
the ultimate way to bring people together. Since then, research
indicates that the effect may be the opposite for many. Social media may
encourage social isolation, and ultimately, social media users may feel
more disconnected from others than those who do not use social media. We
know very little about this phenomenon in China. China presents an
interesting case because the internet is bifurcated to some degree --
there is the Chinese internet and the outside internet available via
VPNs. Being isolation to the outside world may lessen feelings of
isolation, or it may increase these feelings because digital exposure to
a world is much larger than that in people's immediate environment
without the ability to have in-person connection to may make one feel
isolated. Interestingly, and somewhat ironically though, it may also
matter, too, whether someone lives in a city. Research has suggested
that urban living can actually lead to feelings of isolation.

\hypertarget{data}{%
\subsection{Data}\label{data}}

Our data are based on a random Internet survey conducted by Qualtrics
from November 25 to December 2, 2015. We designed the instrument,
content and structure. After we designed the instrument in English and
translated it into Chinese, we entered the questions into the graphical
interface provided from our survey provider, Qualtrics. We were careful
at this stage to make the formatting clear, and to distribute the
questions across numerous pages to prevent overwhelming respondents with
too many questions at once. This helped us achieve a high completion and
response rate. Additionally, we were particularly attentive to the
possibility of question order effects. Many of the questions are
sensitive and could cue responses to subsequent questions. We grouped
and placed questions strategically to limit any such potential for
response bias. Once we settled on the design, Qualtrics administered the
survey in November 2015. They recruit a large pool of respondents for
various survey projects through online advertising on websites such as
local portals, search engines, social networking services, and/or online
shopping services. Panelists who update their profiles at least once
every 6 months are randomly invited to participate in surveys for which
they qualify. Online points are awarded to the corresponding panelists,
which they may accumulate and exchange for cash or various other
country-specific gifts to serve as an incentive for completion. This
number of points is based on the length of the survey. Ours had over
fifty questions, so respondents received a relatively high number of
points.

The data collection had two phases. For the first phase, Qualtrics
collected and submitted to us a trial run of 286 cases. We were able to
check the data for reliability and adjust the instrument before
proceeding with the second phase, the final data collection. We only
made three adjustments, all of which turned out to be important and
central to our analysis. For the trial run, we only asked respondents
about pro-government Internet posters and did not ask them about
potential net-spy or hostile posters. We amended this oversight and it
became central to the analysis that follows. In addition, much of our
theory focuses on the impact of Chinese Internet users circumventing
government filters on the Internet. We initially had a question about
whether citizens had jumped the wall to read sensitive political
information (see Appendix A -- Q25), but decided to add whether they had
done so for entertainment purposes as well (Q26 - to watch foreign
movies, television shows, etc.). Finally, we also included an attention
filter question as part of a battery of institutional trust questions
where respondents were simply told to select the ``None at all''
response (Q13H). If they did not select this response, we could assume
they were not paying attention to the questions, and the survey ended
for these respondents and the data were not collected for them. In the
end, the full sample included a total of 2292 respondents. This sample
size provides for a roughly ±2 margin of error.

While our sample was randomly selected, the sampling frame was based
only on Internet users that Qualtrics can access. Qualtrics recruits
their frame subjects through online advertising on websites such as
local portals, search engines, social networking services, and/or online
shopping services. We do not contend that our sample is representative
of the Chinese population, as the method of obtaining respondents
obviously greatly limits those citizens with little or no access to the
Internet. That said, we are comfortable that our sample is generally
representative of Chinese Internet users. Given that our research focus
centers on Internet effects, it makes sense to have a large sample of
Internet users. A sampling frame including the entire Chinese population
did not make sense for this research. Half of the population do not
regularly access the Internet according to World Bank and China Internet
Network Information Center (CNNIC) data as of late 2015 when we
conducted our survey. Using a sample of the entire population could
potentially limit the degrees of freedom in our models and the marginals
of nuanced measures of digital media use would likely become too small
to have any value. Thus, the large sample of Internet users adds to our
confidence in the inferences throughout the book.

>>>>

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{siunitx}

  \newcolumntype{d}{S[
    input-open-uncertainty=,
    input-close-uncertainty=,
    parse-numbers = false,
    table-align-text-pre=false,
    table-align-text-post=false
  ]}

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The Spatial Relationship between Social Media, Digital Connectedness, and Social Isolation in China: Revisting Cyber-Optimism},
  pdfkeywords={Social Isolation, Social Media, Digital
Communication, China},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{The Spatial Relationship between Social Media, Digital
Connectedness, and Social Isolation in China: Revisting Cyber-Optimism}
\author{}
\date{2023-08-31}

\begin{document}
\maketitle
\begin{abstract}
Early research and popular discourse on the promise of the internet's
impact on social isolation was optimistic. Since these early
predictions, research has become less positive, usually pointing to the
negative role of social media. However, little is known about this
relationship in highly controlled, autocratic social media environments.
Using data from a nationwide survey of Chinese citizens, we do find that
the more citizens digitally connect with others, the less personally
isolated they feel. Conversely, though, the more likely people are to
move relationships from online to offline, and the more they use social
media, the more personally isolated they feel. We also find that this
relationship interacts with urbanicity - being in rural areas tends to
amplify the effect. Overall, our findings suggest that institutional
context may not play an important role in the social media - social
isolation relationship.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, enhanced]}{\end{tcolorbox}}\fi

\newpage{}

@@

Research is quite mixed when it comes to identifying the relationship
between social media use and feelings of isolationism. Early research
suggested that the internet would build social capital, or an
interconnectedness between people, and that social media platforms were
the ultimate way to bring people together. Since then, research
indicates that the effect may be the opposite among at least some part
of the population. Social media may encourage social isolation, and
ultimately, social media users may feel more disconnected from others
than those who do not use social media. We know very little about this
phenomenon in China. China presents an interesting case because the
institutional structures, level of development and the cultural context
are all quite different than the areas studied in most existing studies.
This allows us to test whether such features moderate or exacerbate the
relationship between social media use and social isolation. To fully
test this theory, we employ a large survey of Chinese internet users and
test a variety of regression specifications.\footnote{All data and
  analysis files for this research are available at the author's
  website.} The main result is that social media use is robustly
negatively related to social isolation, though the nature of this
relationship changes depending on the urbanicity of the respondent -
being in a rural area amplifies this relationship.

\hypertarget{literature-review-and-theory}{%
\subsection{Literature Review and
Theory}\label{literature-review-and-theory}}

The existing literature on social isolation and social media use
generally finds a negative relationship between the two. However, there
are some remaining questions regarding whether this relationship remains
significant in different cultural and regulatory environments. Based on
the existing literature, we then develop some testable hypotheses
regarding how the relationship between usage and isolation may vary in
the context of China. The result of these hypotheses tests will help to
better understand the mechanism that links the two variables.

Initially, scholars imagined that social media would create more
opportunities for users to create unique and fulfilling social
relationships. The early theorizing about the possible benefits of
social media hypothesized three important mechanisms by which social
media could decrease loneliness. 1) creating more opportunities for
online to offline meetups and 2) helping to alleviate feelings of
isolation for those who have difficulty making in-person social
connections via joining an online communities, and 3) facilitating
keeping in touch with family and friends (Ellison et al., 2007;
Steinfield et al., 2008; Subrahmanyam et al., 2008).

More recent studies on the subject have found significantly more mixed
and often negative evidence regarding the impact of social media on
feelings of isolation. Primack et al.~found a negative relationship
between social media use and social isolation in youth in a
representative sample (Primack et al., 2017). A meta-review of the
literature on adolescent social media usage notes that while the effect
size is often small the relationship in most studies on social media's
linkage to ill-being is negative (Valkenburg et al., 2022) while
emphasizing the importance of heterogeneous effects. There are
relatively fewer studies that consider all ages. However, those that
have been conducted also emphasize that the relationship between social
media use and well-being has possible negative effects but the effects
seem to vary by personality type (Appel et al., 2020). More recent
research has reinforced the finding that the effect of social media on
measures like social isolation may be highly heterogeneous across
individuals; for some individuals, the effect may be quite negative for
some while for others social media usage may have a modestly positive
impact on well-being (Beyens et al., 2020; Ostic et al., 2021).

While most of the studies agree that the causal direction flows from
social media use to feelings of isolation (though see (Kim, 2017;
Nowland et al., 2018)), there is disagreement among authors about the
specific mechanisms and causal pathways through which this relationship
exists. For those who find a negative relationship between social media
and well-being, one set of mechanisms suggests that the social media
itself that is the problem. Cyberbullying, or more generally, repeated
and aggressive online attacks against users has been systematically
linked to feelings of social isolation (Ademiluyi et al., 2022). Another
proposed mechanism is upward comparisons of others and the weak nature
of online connections - both of these have been found to generate social
anxiety and social isolation (Alabri, 2022; Büttner and Rudert, 2022). A
third set of mechanisms primarily concern themselves with the physical
changes that social media causes. Social media use has been shown to
displace face to face activity, which leads to greater social isolation
(Larson et al., 2018). Heavy social media use also decreases hours of
sleep, leading to greater feelings of anxiety and more social isolation
(Twenge et al., 2017). The current research suggests that which of these
mechanisms is dominant depends on the age/demographic group in question
and the societal context (Twenge, 2019).

Most of the existing research on the subject focuses on Western highly
developed societies, and this focus may limit variation on key
independent variables given the similar social and institutional context
in which users are embedded. To address this problem, this study aims to
understand how these complex dynamics interact in the non-Western
setting of China. Relatively little is known about China's level of
social isolation, though what evidence that does exist suggests that it
is on the rise. Twenge et al.'s study found that adolescent feelings of
loneliness have generally been on the rise in comparable East Asian
countries (Twenge et al., 2021). News reports have highlighted the
increasing number of people who consider themselves to be online
outcasts or losers (\emph{diaosi}) (China{'}s latest social media trend:
Diaosi / loser subculture, 2016; Nan, 2013; Szablewicz, 2014). We do not
aim to make a direct comparison between China and Western countries but
rather highlight the way in which China's unique context helps shed
light on which contextual factors matter in understanding the linkage
between the two variables.

The most important of these contextual factor in China is that social
media space that is highly controlled online environment. Many people
are aware that the Chinese government heavily censors politically
sensitive content on online platforms. However, fewer are aware that the
Chinese government spends as much, if not more, effort to clean up
non-political as negative or distressing content on the Chinese
internet, such as scams, instances of cyber bullying, disturbing videos,
and vulgar content (Wang, 2020). This control impulse also extends
content the government views as spiritually polluting, such as
pornography, scantily dressed women, and other escapist content. While
the Chinese government and online platforms are not 100 per cent
effective in their ability to completely stamp out such content, the
default is that such content is actively restricted, while US platforms
tent to take a more pro-speech regulatory position (Liu and Yang, 2022).
If one of the main mechanisms through which social media promotes social
isolation occurs is exposure to negative content, then this highly
regulated online environment should lead to a weaker relationship
between social media use and isolation. Suggestive of this is a a study
by Liu and Liu that finds Chinese who consumed more commercialized and
sensationalized media reported more trauma during the first COVID-19
outbreak in China than those who followed government sources (Liu and
Liu, 2020).

On the other hand, China also has a highly robust online influencer
culture. While social media influencers are a global phenomena, China
has perhaps the most commercialized version of influencing. Social media
influencers are developed by large agencies and work in an almost
assembly-line fashion to promote products on livestreams or pitch
sponsor products. In this influencer culture, less attention is paid to
authenticity and personality and relatively more emphasis is placed on
consumerism, brand promotion, and image (Tam, 2019; Wei, 2023). Given
that previous research has suggested one of the main channels through
which social media creates feelings of isolation is negative personal
comparisons with online content (so-called ``FOMO'' - fear of missing
out), heavy social media use may increase the strength of the
relationship between social media use and social isolation through
consumption of such highly commercialized content.

Additionally, China's social media environment is relatively unique in
that the dominant messaging platform is also the largest social media
app (WeChat). Users can use it simply to chat with family and friends,
they can use it to post photos and other content about their life in a
feed that works similarly to Facebook, or they can browse videos of
others in a method similar to Tiktok. Previous research has indicated
that while, on the whole, social media increases social isolation, for
users that primarily use it to contact friends and family, its use can
counteract the negative effects (Hajek and König, 2019; Lim et al.,
2020). In this way, those that heavily use these type of combined
platforms may engage in multiple uses of social media that tend to
cancel each other out. Given the ubiquity of WeChat in China, such a
relationship seems plausible.

Finally, the nexus of social isolation and social media use in China has
played out in the context of one of the most rapid periods of
urbanization in human history. Hundreds of millions of people have
shifted from rural subsistence farming to participating in an urban
market economy. The existing literature finds that urbanicity is an
important factor in levels and mechanisms for social isolation; many
studies of Western societies have found that rural residents, while
having stronger familial bonds, also can more easily experience social
isolation without those bonds (Henning-Smith et al., 2018; Kaye, 2017;
Koning et al., 2017). In China, less is known about the differential
effects of the urban rural divide on feelings of social isolation but
the research that does exist suggests that urban environments can be
especially isolating, particularly for rural to urban migrants
particularly due to lack of access to traditional support systems and
familial connections (Li and Rose, 2017).

Given the existing literature and the context of China, this leads to a
series of testable hypotheses.

\begin{itemize}
\item
  \textbf{H1a:} social media use should have a weak or possibly positive
  connection to social isolation due to strengthened government control
  over online content
\item
  \textbf{H1b:} social media use should have a negative connection to
  social isolation due to the dominance of influencer culture in China
\item
  \textbf{H2:} those who are able to use social media to make new
  friends will have lower social isolation scores than those who do not
\item
  \textbf{H3:} the urban/rural divide will produce differential impacts
  in the above hypotheses on the relationship between social media use
  and social isolation
\end{itemize}

\hypertarget{data-measurement-and-descriptives}{%
\subsection{Data, Measurement, and
Descriptives}\label{data-measurement-and-descriptives}}

The data used in this study are original. We designed a survey
instrument in English, and translated it into Chinese, to measure a
range of concepts including social isolationism and various dimensions
of social media and digital information consumption. We employed
Qualtrics to collect the data. They randomly selected 2292 respondents
from their existing panel\footnote{Qualtrics recruits a large pool of
  respondents for various survey projects through online advertising.
  Recruits who update their profiles at least once every 6 months are
  randomly invited to participate. These recruits are awarded online
  points that can be exchanged for cash or various other
  country-specific gifts. The number of points is based on the length of
  the survey, and since our survey had over fifty questions, respondents
  received a relatively high number of points.} from November 25 to
December 2, 2015. This sample size on provides for a roughly \(\pm2\)
margin of error.\footnote{We were able to first collect a small sample
  (\(n=286\)) to check the reliability and adjust the instrument after
  before proceeding with the final data collection. We made three
  adjustments, none of which are related to the measures used in the
  current study.}

Our dependent variable, \emph{social isolationism}, is a three-item
additive index. Respondents were given the following introduction: The
next question is about how you feel about different aspects of your
life. Could you tell me for each one if you feel that way always, almost
always, some of the time, rarely or never? Then they were asked: A) How
often do you feel that you lack companionship, B) How often do you feel
isolated from others?, and C) How often do you feel left out? Response
options were recoded so that higher responses equated to feeling more
socially isolated, they were all three added together, and then rescaled
to range from 0 through 1 maintaining the original intervals
(\(\alpha = 0.86\)). The distribution of that index is presented in
Figure~\ref{fig-social-iso-dist}. The distribution is relatively normal
centered around the midpoint with a slight skew to the right. Most
respondents appear to be on the low end, indicating the the majority of
those in our sample do not feel particularly isolated. Conversely, the
right skew in the distribution does suggest that a sizable chunk of
folks feel quite socially isolated. Altogether, the spread out
distribution indicates quite a bit of variance (\(\mu = 0.41\),
\(sd = 0.20\)).

\begin{figure}

{\centering \includegraphics{Social-Isolation-in-China_files/figure-pdf/fig-social-iso-dist-1.pdf}

}

\caption{\label{fig-social-iso-dist}Distribution of Social Isolationism
Index}

\end{figure}

We have several primary independent variables including human
interaction, online to offline relationships, general social media use,
and urbanicity. We measured human interaction with a single item: ``Has
the internet and phone applications increased your contact with the
following groups of people (check yes to all that apply)?'' The response
options were: 1) Family that lives nearby, 2) Family that lives far
away, 3) Friends that live nearby, 4) Friends that live far away, 5)
People you met on the internet that live nearby, and 6) People you met
on the internet that live far away. We simply counted the number of
selected responses so the item ranged from 0 to 6, and then for the
models that come later, we rescaled it to range from 0 through 1
maintaining the original intervals. Online to offline relationships was
also measured with a single item: Have you met someone offline that you
initially met online? Response options were: 1) Yes, many times, 2) Yes,
several times, 3) Yes, once, and 4) No, never. We inverted this scale so
that higher values represented more offline relationships, and again,
rescaled it to range from 0 through 1 for the models that follow.

The distributions of these ordinal independent variables are reflected
in Figure~\ref{fig-var-dist}. Very few people claimed to never have
increased their contact/connections with others through the internet.
The modal number of increased connections respondents claimed to have
made through the internet was 2, followed closely by 1, but a sizable
proportion of the sample, about 48\% claimed to have increased their
number of personal connections by 3 or more through the internet.
Perhaps not surprisingly, the modal response when asked whether one had
met someone online and that relationship moved offline was ``No,
never''. On the other hand, we were surprised that about 37\% of those
sampled claimed to have moved relationships offline several times, over
10\% said they had done so many times, and about 13\% did so once. Taken
altogether, a large proportion of our respondents seem to using the
internet to facilitate social relations.

\begin{figure}

{\centering \includegraphics{Social-Isolation-in-China_files/figure-pdf/fig-var-dist-1.pdf}

}

\caption{\label{fig-var-dist}Primary Independent Variables
Distributions}

\end{figure}

We measured general social media use with an index based on the
following three items: 1) About how many hours a day would you estimate
you spend using only social media? Social media means applications like
Weibo, QQ, Renren, Kaixin001, Douban, WeChat or other sites and services
that allow users to interact with each other. (0-1, 1-2, 2-3, 3-4, 4-5,
5-6, 6-7, 7-8, 8-9, or More than 9), 2) Do you check email, read
websites, and use social media (social media means applications like
Weibo, QQ, Renren, Kaixin001, Douban, WeChat or other sites and services
that allow users to interact with each other) more than you did five
years ago? (Yes, No), and 3) How often do you read news stories about
political events that have been posted on social media (social media
means applications like Weibo, QQ, Renren, Kaixin001, Douban, WeChat or
other sites and services that allow users to interact with each other)?
(More than once a day, Everyday, Three-to-five days per week, One-to-two
days per week, Less often, Never). Each was recoded so that higher
values represented more social media use, rescaled to range from 0
through 1, added together, and then again, rescaled to range from 0
through 1 maintaining all original intervals. The distribution of this
index is presented in Figure~\ref{fig-sm-use-dist}. The distribution is
relatively normal and tightly centered around the mean 0.63 with a
standard deviation of 0.17, but there is a left skew suggesting that a
portion of the population is not that social media active, but clearly
the bulk of the population are.

\begin{figure}

{\centering \includegraphics{Social-Isolation-in-China_files/figure-pdf/fig-sm-use-dist-1.pdf}

}

\caption{\label{fig-sm-use-dist}General Social Media Use Distribution}

\end{figure}

Our final primary independent variable, urbanicity, was measured using a
single item: Which of these best describes the place in which you live?
(Countryside/Village, Small City, Mid-Sized City, Suburban Area of a Big
City, Big City). For the purpose of the models, we rescaled this item to
range from 0 through 1 maintaining the original intervals. The
distribution is reflected in Figure~\ref{fig-urban-dist}. Given the
rural to urban Chinese migration since the founding of the People's
Republic (Xia, 1995), it is not surprising to see that nearly 45\% of
those in our sample live in cities, 7\% in suburbs, and about 27\% in
mid-sized cities. About 16\% say they live in small cities and only
about 5\% say they live in villages.

\begin{figure}

{\centering \includegraphics{Social-Isolation-in-China_files/figure-pdf/fig-urban-dist-1.pdf}

}

\caption{\label{fig-urban-dist}Urbanicity Distribution}

\end{figure}
